---
title: "Individual1"
output: html_document
author: Sam Shifflett
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
bio <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
```

Cleaning Data
1)
```{r}
defense[,c(1,2)] <- sapply(defense[,c(1,2)], as.factor)
defense[,3:11] <- sapply(defense[,3:11], as.numeric)
defClean <- defense
offense[,c(1,2)] <- sapply(offense[,c(1,2)], as.factor)
offense[,3:12] <- sapply(offense[,3:12], as.numeric)
offClean <- offense
bio[,c(3,4)] <- sapply(bio[,c(3,4)], as.numeric)
bioClean <- bio
```

Tidying(1)
1)
```{r}
library("tidyverse")
defGather <- gather(defClean,Statistic,Value,-c(1,2))
defGather
```

2)
```{r}
p <- ggplot(defGather, aes(x = Opponent_Opponent, y=Value)) + geom_histogram(stat="identity") + facet_wrap(~Statistic) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab("Opponent") + ylab("Total Amount") + ggtitle("Defensive stat count per game")
p
```

The Turnovers are the most rare occurences in games so causing turnovers is a special skill.

3)
```{r}
defCleanFiltered <- defClean %>% filter(Opponent_Opponent == "West Virginia" | Opponent_Opponent == "Kansas State")
defCleanSolo <- defCleanFiltered[,1:3]
defCleanSolo <- defCleanSolo %>% filter(Tackles_Solo != 0)
df <- defCleanSolo %>% spread(key=Opponent_Opponent, value=Tackles_Solo)
df[is.na(df)] <- 0
names(df) <- c("Player","KS","WV")
ggplot(df,aes(x=WV,y=KS)) + geom_point() + geom_smooth(method="lm") + ylim(0,10) + xlim(0,10)
```

Though the trendline shows a slope favoring the West Virginia game, the total amount of tackles in the Kansas state game was more so ISU had a better Defensive game against Kansas State.


Tidying(2)
1)
```{r}
dfBio <- bioClean %>% separate(Hometown,c("City","State"),sep=",")
eachState <- dfBio %>%
  group_by(State) %>%
  summarise(totalPlayer = n())
head(eachState)
```

```{r}
mergeDf <- merge(bioClean,offClean)
recYds <- mergeDf %>% group_by(Name) %>% summarise(totalRYDS = sum(Receiving_YDS))
weight <- mergeDf[,c(1,4)]
singleWeight <- unique(weight)
df1 <- merge(recYds,singleWeight)
ggplot(df1,aes(x=totalRYDS,y=Weight)) + geom_point() + geom_smooth(method="lm")
```

There doesn't seem to be much of a correlation between the weight of a player and their total receiving yards. The only thing that I would point out is that there are occurences of heavier players(240lbs or above) receiving in short distances, but none in long distance.


```{r}
df2 <- merge(bioClean,offClean)
df2 <- df2[df2$Position == "Wide Receiver",]
df2 %>% group_by(Class) %>% summarize(totalRYDS = sum(Receiving_YDS))
```

I wanted to see which classes had the most receiving yards. It semms like redshirting is the way to go if you want to be a successful receiver at ISU.


